Percussive Sound Generation with Virtual Listeners and Modular Synthesizers

### 2020-01-31 paper.tex
Updates from Overleaf

\section{Introduction}
\subsection{Terminology}
\todo[inline,color=green!30]{I'd like to define terms such as VSTs, audio effects, Synthesizers, paramters and presets here. I throw these terms around a lot and I think some might be confused by their meaning. Is this the appropriate place for that? I'd like the reader to know these definitions before reading the intro. I can try to keep it brief and engaging}
\subsection{Background and Motivation}
The rise of Digital Audio Workstations (DAW) \cite{leider2004digital} and Virtual Studio Technology (VST) based plug-ins \cite{tanev2013virtual} have rapidly transformed the sonic and material landscape of music production in the recent years. Coupled with this rise in popularity is a vast array of commercial products and services dedicated to satiating the need of amateur and professional music producers for unique sounds; most commonly via audio samples: one-shot drum samples, long sustained notes (commonly referred to as pads or textures), and loops (percussive or melodic) are common deliverables. Two notable examples of these commercial services are \textit{loopmasters}\footnote{loopmasters.com} and \textit{splice.com}\footnote{splice.com}. Furthermore, VST plug-ins can emulate complex audio synthesizers and effects which some producers may find daunting or time consuming to work with from scratch. In many cases VST plug-in vendors or unaffiliated enthusiasts sell additional presets for these plugins, targeted towards producers who do not have the time or interest in creating their own. Producers may further modify these presets until their desired sound is reached.\\
Our work is motivated by the idea of finding new, convenient methods for the expansion of a music producer's library of sounds. Primarily with generation of novel, one-shot audio samples but also with automated search and creation of presets for virtual synthesizers that a producer might find  interesting. Using the generation of short, percussive audio samples as a starting point this project is a proof of concept for promising avenues towards our motivational goal.\\

\subsection{Methodology}
Towards our goal of generating original audio, we found the proper implementation of 2 major components to be crucial:
\begin{itemize}
    \item \textit{Virtual Synthesizer}: A flexible, deterministic, and tractable generator which can create audio. The core of our implementation made use of pippi, \footnote{https://github.com/luvsound/pippi} a fast, offline focused python DSP library with a C backend. We additionally used the scipy\cite{jones2001scipy} library a few audio effects that we found lacking in pippi. 
    \item \textit{Virtual Ear}: An ear that returns an evaluation of an audio sample; estimating the effectiveness of an audio sample's fulfillment of a producers requirements. The ear's evaluation guides the generation process towards a desired path, making the implementation and training of the ear crucial to the sounds which are generated. A particularly effective implementation was done by using a labeled set of percussive and non-percussive audio files to train a custom model (combination of a simple Convolutional Neural Net (CNN) and a custom envelope estimation algorithm), giving us a "virtual drum detector ear". \\
\end{itemize}
Our components are designed with modularity and parallelizability in mind. This allows each component to be debugged, modified, and improved without requiring modifications in other components; additionally increasing the scalability and speed of experiments.
Section \ref{impl} contains further discussion of the components as well as the code that glues the project together.\\ 
While the main focus of this project is the generation of novel percussive sounds our methodology indicates promising results with regards to creation of new presets for any virtual synth without the need for a-priori knowledge of the functions or its parameters (i.e the effect of parameter modulation on the sonic output). We also demonstrate the viability of a few lesser explored avenues for the purposes of audio synthesis and Music Information Retrieval (MIR), notably:
\begin{enumerate}[label=\roman*]
\item Heuristic search methods for the creation of new presets for virtual synthesizers. More concretely, we observed that given a virtual ear which could reliably score pieces of audio based on its resemblance to a desired category (e.g: kicks, snares, piano) we were able to rapidly search the parameters of a virtual synthesizer and create numerous parameter-sets (i.e presets) which resemble our desired category of sounds.
\item The viability of virtual synthesizers based on Digital Signal Processing (DSP) methods for fast, unsupervised creation of novel audio, discussed in more detail in section \ref{related}. \\
\end{enumerate}

\subsection{Related Work and Contradistinctions}
\label{related}
Numerous deep, neural network models have been proposed and utilized for the purpose of signal generation in recent years. WaveGans and WaveNet have been subject to significant improvements and experiments since their proposal \colorbox{green!=40}{\cite{nsynth2017} + Many more references to be added.} Even more recently Variational AutoEncoders (VAE's) have been utilized for generation of short percussive samples \cite{aouameur2019neural,ramires2019timbfeat}. In this work however, we opt to use digital signal processing methods to create a virtual synthesizer for the generation of audio signals as it provides several unique advantages:
\begin{enumerate}[label=\roman*]
  \item Fast, offline rendering of audio with no reliance on GPU: Currently not possible with state of the art models such as parallel WaveGan \cite{yamamoto2019parallel} and parallel WaveNet \cite{oord2017parallel}. 
  \item Rendering at high sampling rates: Performance speed being a common issue, the standard sampling rate in most audio generation work utilizing neural networks appears to be under 24 khz \cite{yamamoto2019parallel,oord2017parallel,aouameur2019neural,ramires2019timbfeat}. However, a significant number of untrained human ears can detect a change in quality of audio between sampling rates of 192 khz and the industry standard of 44.1 khz \cite{reiss2016meta} with a dramatic increase in quality detection after training. Therefore we can safely assume that most producers would prefer their audio samples to have sampling rates of 44.1 khz or higher. 
  \item Neural networks are often viewed as unexplainable black box solutions. Some models such as VAE's can learn an underlying latent space of parameters and capture the "essence" of the different labels in a dataset. However, these spaces are learned in an unsupervised manner and must be manually analysed, perhaps extensively, before they can be understood \cite{esling2018generative}. The use of a virtual synthesizer for audio generation makes our parameters readily understandable and easily modifiable. \\

\end{enumerate}
Automatic programming of virtual synthesizers has also been a topic of interest. In early 2000s, Interactive Genetic Algorithms (IGA's) were utilized for the generation of new sounds with various sound-engines \cite{johnson1999exploring,dahlstedt2001creating}. More recent work by Yee-King et al. \cite{yee2018automatic} used LSTM and genetic algorithms to find the exact parameters used to create a group of sounds. The sounds approximated were made by the same virtual synthesizer, not an external source; making the eventual replication certain even with random search. Since this work more focused on pads and textures rather than drums, feature matching in this also appears to not be concerned with the envelope of the sounds but rather the frequency content within arbitrary time windows. Yet another recent, impressive work by Esling et al. used a large dataset of over 10,000 presets for a commercial VST synthesizer to learn a latent parameter space which can be sampled for creation of new audio \cite{esling2019universal}. As stated before, our work presents a rapid approximation of percussion sounds with no previous knowledge about the sonic capabilities of our virtual synthesizer, exploring the actual parameter space rather than its latent representation.

\subsection{Data And Project Replication}
\label{data}
Our data is a large set of drum samples aggregated from personal libraries, free drum kits from sampleswap \footnote{https://sampleswap.org/} which we further processed to suit our categories, and a large set of drum sounds aggregated from royalty free sources such as musicradar \footnote{https://www.musicradar.com/}. We will make our dataset of free-drum sounds available for download. The scripts used to download and process royalty free samples will also be made available. Further information about downloading our dataset can be found on the project's github page. \\
Our drum categories are claps, hats, kicks, rims/other, shakers, snares, mid-toms and high-toms. Other categories are chopped guitars, chopped pianos and stacks (randomly generated synth noise), utilized for learning percussive vs not-percussive sounds. Some notes about our dataset:
\begin{itemize}
\item Of the 6000 drum-sounds utilized in our work, the kick, snare and hat categories have the largest share at around 20\% each, while the shaker and rim (other) categories have the smallest at 5\% combined. Due to this we only focused on learning from kicks, snares, toms, claps and hats for Phase 1 of training \ref{sec:ear}.
\item For Phase 2 of training we only focused on categorizing snares, claps, kicks, hats and other (percussive sounds such as shakers, rims and unusual percussions that we couldn't categorize were grouped into this category). 
\item In order to offset bias from data imbalance during training of our models, the categorical cross entropy loss was weighted by the category sizes. 
\item For training any given model, 80\% of our data is used for training the models and 20\% is used for testing. 
\item We limit the size of the stack category to 50\% of the total size of our drum dataset. This is done in order to measure whether the features extracted can address the "Open Set Recognition" problem, which will be discussed further in section \ref{sec:ear}.
\end{itemize}
\section{Implementation}
\label{impl}
\subsection{Virtual Synthesizer}
We used the python based Pippi library for sound generation\footnote{https://github.com/luvsound/pippi}. This library uses a C back-end and focuses on fast offline generation of audio signals. In our implementation a synthesizer can have any number of sub-modules. The parameters that deterministically dictate the output signal of the sub-module as well as the range of values each parameter can take are shown in table 1 \ref{table:1}. We call this number the \textit{stack size}. For the range of parameters each sub-module can take we call the sets of parameters that characterize a synth's sub-modules a \textit{program} (analogous to a preset).  Each sub-module can make an audio signal with the length of 0.1-1 second. If a Synth has a stack size of more than 1 these audio signals are overlapped and the total amplitude is normalized.\\
\begin{table}[h!]
\centering
\resizebox{\columnwidth}{!}{\begin{tabular}{ |c|c|c| } 
\hline
Parameters & Value Range & notes and constraints\\
\hline \hline
Attack & 0-3 & A-D-S-R values relative\\
Decay & 0-3 & relative to A-S-R\\
Sustain & 0-3 & relative to A-D-R\\
Release & 0-3 & relative to A-D-S\\
OSC type & sine,square,saw & tone type\\
IsNoise & boolean & whether to generate noise\\
Sample Length & 0-1 second & - \\
StartTime & 0-1 second & Length+Start<1\\
Amplitude & 0.1-1 & 1 = max amplitude\\
Pitches(notes) & list of pitches &  range of C0(16.35hz) to B9 \\
HP filter Cuttoff & 0-20000 & -\\
LP filter Cuttoff & 20000-HP & never lower than HP cutoff\\
Filter Order & 4,8,16 & butterworth filter order \\
\hline
\end{tabular}}
\caption{Synthesizer Sub-module Parameters. Despite the simplicity of the parameters and our efforts at constraining the ranges, the number of parameters that can be randomly chosen for each sub-module is in the order of $10^{15}$ }
\label{table:1}
\end{table}
\begin{figure*}[h]
\centering

\begin{subfigure}[b]{\linewidth}
\includegraphics[width=1\linewidth]{images/ff1.pdf}
\caption{Organic hat sample}\label{fig:64stack}
\setcounter{subfigure}{1}%
\includegraphics[width=1\linewidth]{images/ff2.pdf}
\caption{Randomly generated audio with percussive qualities, resembling a tight snare}\label{fig:64stack}
\setcounter{subfigure}{2}%
\includegraphics[width=1\linewidth]{images/ff3.pdf}
\caption{A randomly generated noise with a percussive envelop but non-percussive frequency features (modulated pitch)}\label{fig:64stack}
\setcounter{subfigure}{2}%
\end{subfigure}

@@

What we refer to as an "ear" is any method of evaluating a piece of audio and capable of "listening" to a piece of audio and giving it a score (or a list of scores) based on how well it satisfies certain criteria. Since in this work we are mainly focused on percussive generation what we require from the ear is to give us probabilities of an audio sample belonging to various categories. As our synthesizer outputs are deterministic for all programs the ear's evaluations would allow us to associate the categorical probabilities of each sound with the program that generated it. In section \ref{gens} we discuss how these scores were used for navigation of our synthesis towards parameters which give us the desired sonic output.\\

@@

 \subsubsection{Survey of Randomly Generated Drums}
\label{survey}
\begin{figure}[H]
\centering
\includegraphics[width=1.1\linewidth]{images/cat.pdf}
\caption{Frequency of assigned labels for the human categorizers and the average categorizer}
\label{fig:freq-survey}
\end{figure}
To measure the quality of the samples produced by our pipeline and the power of our models, we randomly generated around 50-60 samples in the following categories: "snare", "kick", "hat", "clap" and "other" (combination of rims, 
shakers and unusual percussive sounds"). These samples were determined to be percussive and then categorized by 4 different models (FC, CNNLSTM, E+F and AVG). We ensured a balanced division between samples of stack size 1, 2 and 4 (each stack is responsible for a third of the samples under each category). Both writers then categorized these samples without knowledge of other categorizations (human or computation models). It's important to note that:
\begin{itemize}
    \item Humans had an additional category of "bad" for samples that they deemed not percussive. The "bad" category indicates that the sample should have been skipped in Phase 1. 
    \item With 6 categorization groups, humans had the same categorization in 47\% of cases.
    \item The agreement between the humans and AVG was 44\% and 47\%, not significantly lower than agreement with each other. 
    \item Of 257 samples humans agreed with FC, CNNLSTM, AVG and E+F respectively in 78, 76, 76 and 46 of cases.
\end{itemize}
We assess the reliability of agreement between humans and categorization models via the Fleiss' kappa coefficient \cite{fleiss1971measuring}. The value of 0 or less for this coefficient indicates no agreement beyond random chance, and the value of 1 indicating perfect agreement. Our kappa measurements \ref{kappa_table} lie within the 3.5-4.5 range, indicating mild to moderate agreement between humans and machines. We again measure this coefficient after dropping samples that were categorized as "bad" by the authors, as samples that humans deem to be "bad" indicate a failure in Phase 1 and arguably should not have been categorized by the models at all. Dropping of samples that both authors deemed "bad" causes an 8\% reduction of our data (21 samples) and a small increase in kappa score. Dropping samples either reviewer deemed bad resulted in a 30\% reduction of samples and relatively large increase in kappa scores. \\

@@

\begin{itemize}
    \item Our methodology was successful in generating novel examples of some percussion categories.
    \item Reiterate the importance of ear.
    \item There are potential benefits (e.g: pedagogical, software engineering, efficiency, data search, generality of solution, etc.) to separate some audio generation tasks into multiple parts, rather than throwing everything at some neural net. 
    \item With the right implementation of DSP methods to create a synth, Genetic search (viewed as too slow by previous work) is viable for finding desired programs and parameters.
    \item Neural nets are function approximators. You don't need them if the actual function is understandable and simple...(but is our task simple enough? how do I prove it?)
    \item When possible, using DSP methods instead of NNs is preferable in many ways (speed, computation, hardware requirements).\\
    \colorbox{green!=30}{Include some speed measurements in appendix?}
    \item ...
\end{itemize}
\section{Future Work}
\begin{itemize}
    \item Potential for other generative methods (e.g reinforcement learning)
    \item Making the synth's sound range more diverse while maintaining speed and efficiency
    \item Can our approach be used for VSTs?
\end{itemize}

>>>>

\section{Introduction}

\subsection{Background and Motivation}
The rise of Digital Audio Workstations (DAW) \cite{leider2004digital} and Virtual Studio Technology (VST) based plug-ins \cite{tanev2013virtual} have rapidly transformed the sonic and material landscape of music production in the recent years. Coupled with this rise in popularity is a vast array of commercial products and services dedicated to satiating the need of amateur and professional music producers for unique sounds; most commonly via audio samples: one-shot drum samples, long sustained notes (commonly referred to as pads or textures), and loops (percussive or melodic) are common deliverables. Two notable examples of these commercial services are \textit{loopmasters}\footnote{loopmasters.com} and \textit{splice.com}\footnote{splice.com}. Furthermore, VST plug-ins can emulate complex audio synthesizers and effects which some producers may find daunting or time consuming to work with from scratch. In many cases VST plug-in vendors or unaffiliated enthusiasts sell additional presets for these plugins, targeted towards producers who do not have the time or interest in creating their own. The flexibility of the VST technology allows producers to further modify these presets until their desired sound is reached.\\
Our work is motivated by the idea of finding new, convenient methods for the expansion of a music producer's library of sounds. Primarily with generation of novel, one-shot audio samples but also with automated search and creation of presets for virtual synthesizers. Using the generation of short, percussive audio samples as a starting point, this project is a proof of concept for a promising avenue towards our motivational goal.\\

\subsection{Methodology}
Our attempt is centered around a generateive pipeline of audio. We want random noises to be created by a tractable source, we also want to evaluate these sounds so we can guide the random noise generator towards desired sounds. Following this idea, we found the proper implementation of 2 major components to be crucial:
\begin{itemize}
    \item \textit{Virtual Synthesizer}: A flexible, deterministic, and tractable generator which can create audio. 
    % The core of our implementation made use of pippi, \footnote{https://github.com/luvsound/pippi} a fast, offline focused python DSP library with a C backend. We additionally used the scipy\cite{jones2001scipy} library a few audio effects that we found lacking in pippi. 
    \item \textit{Virtual Ear}: An ear that returns an evaluation of an audio sample; estimating the effectiveness of an audio sample's fulfillment of a producers requirements. The ear's evaluation guides the generation process towards a desired path, making it a crucial component of our pipeline. 
\end{itemize}
Our components are designed with modularity and parallelizability in mind. This allows each component to be debugged, modified, and improved without requiring modifications in other components while additionally increasing the scalability and speed of experiments. 
Section \ref{impl} contains further discussion of the components as well as the code that glues the project together.\\ 
While the main focus of this project is the generation of novel percussive sounds, our methodology indicates promising results with regards to creation of new presets for any virtual synth without the need for a-priori knowledge of the functions or its parameters (i.e the effect of parameter modulation on the sonic output). We also demonstrate the viability of a virtual synthesizers based on Digital Signal Processing (DSP) methods for fast, unsupervised creation of novel audio, discussed in more detail in section \ref{related}. \\
\subsection{Related Work and Contradistinctions}
\label{related}
Numerous deep, neural network models have been proposed and utilized for the purpose of signal generation in recent years. WaveGans and WaveNet have been subject to significant improvements and experiments since their proposal \cite{nsynth2017,yamamoto2019parallel,oord2017parallel}. Even more recently Variational AutoEncoders (VAE's) have been utilized for generation of short percussive samples \cite{aouameur2019neural,ramires2019timbfeat}. In this work however, we opt to use digital signal processing methods to create a virtual synthesizer for the generation of audio signals as it provides several unique advantages:
\begin{enumerate}[label=\roman*]
  \item Fast, offline rendering of audio with no reliance on GPU: Currently not possible with state of the art models such as parallel WaveGan \cite{yamamoto2019parallel} and parallel WaveNet \cite{oord2017parallel}. 
  \item Rendering at high sampling rates: Performance speed being a common issue, the standard sampling rate in most audio generation work utilizing neural networks appears to be under 24 khz \cite{yamamoto2019parallel,oord2017parallel,aouameur2019neural,ramires2019timbfeat}. However, a significant number of untrained human ears can detect a change in quality of audio between sampling rates of 192 khz and the industry standard of 44.1 khz \cite{reiss2016meta} with a dramatic increase in quality detection after training. Therefore we can safely assume that most producers would prefer their audio samples to have sampling rates of 44.1 khz or higher. In this work, we fix our sampling rate to the 48 khz standard. 
  \item Neural networks are often viewed as unexplainable black box solutions. Some models such as VAE's can learn an underlying latent space of parameters and capture the "essence" of the different labels in a dataset. However, these spaces are learned in an unsupervised manner and must be manually analysed, perhaps extensively, before they can be understood \cite{esling2018generative}. The use of a virtual synthesizer for audio generation makes our parameters readily understandable and easily modifiable. \\

\end{enumerate}
Automatic programming of virtual synthesizers has also been a topic of interest. In early 2000s, Interactive Genetic Algorithms (IGA's) were utilized for the generation of new sounds with various sound-engines \cite{johnson1999exploring,dahlstedt2001creating}. More recent work by Yee-King et al. \cite{yee2018automatic} used Long short-Term Memory (LSTM) models and genetic algorithms to find the exact parameters used to create a group of sounds. The sounds approximated were made by the same virtual synthesizer, not an external source; making the eventual replication certain even with random search. Since this work more focused on pads and textures rather than drums, feature matching in this also appears to not be concerned with the envelope of the sounds but rather the frequency content within arbitrary time windows. Yet another recent, impressive work by Esling et al. used a large dataset of over 10,000 presets for a commercial VST synthesizer to learn a latent parameter space which can be sampled for creation of new audio \cite{esling2019universal}. As stated before, our work explores the rapid approximation of percussion sounds with no previous knowledge about the sonic capabilities of our virtual synthesizer, exploring the actual parameter space rather than its latent representation.

\subsection{Data And Project Replication}
\label{data}
Our data is a large set of drum samples aggregated from personal libraries, free drum kits from the sample-swap project \footnote{https://sampleswap.org/} which we further processed to suit our categories, and a large set of drum sounds aggregated from royalty free sources such as musicradar \footnote{https://www.musicradar.com/}. We will make our dataset of free-drum sounds available for download. The scripts used to download and process royalty free samples will also be made available. Further information about downloading our dataset can be found on the project's github page. \\
Our drum categories are claps, hats, kicks, rims/other, shakers, snares, mid-toms and high-toms. Other categories are chopped guitars, chopped pianos and stacks (randomly generated synth noise), utilized for learning percussive vs not-percussive sounds. Some notes about our dataset:
\begin{itemize}
\item Of the 6000 drum-sounds utilized in our work, the kick, snare and hat categories have the largest share at around 20\% each, while the shaker and rim (other) categories have the smallest at 5\% combined. Due to this we only focused on learning from kicks, snares, toms, claps and hats for Phase 1 of training (along with non-percussive groups of sound) \ref{sec:ear}.
\item For Phase 2 of training we only focused on categorizing snares, claps, kicks, hats and other (percussive sounds such as shakers, rims and unusual percussions that we couldn't categorize were grouped into this category). Non-percussive datasets were not used for this phase. 
\item In order to offset bias from data imbalance during training of our models, the categorical cross entropy loss was weighted by the group sizes. 
\item For training any given model, 80\% of our data is used for training the models and 20\% is used for testing. 
\item We limit the size of the stack category to 50\% of the total size of our drum dataset. This is done in order to measure whether the features extracted can address the "Open Set Recognition" problem, which will be discussed further in section \ref{sec:ear}.
\end{itemize}
\section{Implementation}
\label{impl}
\subsection{Virtual Synthesizer}
We used the python based Pippi library for sound generation\footnote{https://github.com/luvsound/pippi}. This library uses a C back-end and focuses on fast offline generation of audio signals. In our implementation a synthesizer can have any number of sub-modules. The parameters that deterministically dictate the output signal of the sub-module as well as the range of values each parameter can take are shown in table \ref{table:1}. We call this number the \textit{stack size}. For the range of parameters each sub-module can take we call the sets of parameters that characterize a synth's sub-modules a \textit{program} (analogous to a preset).  Each sub-module can make an audio signal with the length of 0.1-1 second. If a Synth has a stack size of more than 1 these audio signals are overlapped and the total amplitude is normalized.\\
\begin{table}[h!]
\centering
\resizebox{\columnwidth}{!}{\begin{tabular}{ |c|c|c| } 
\hline
Parameters & Value Range & notes and constraints\\
\hline \hline
Attack & 0-3 & A-D-S-R values relative\\
Decay & 0-3 & relative to A-S-R\\
Sustain & 0-3 & relative to A-D-R\\
Release & 0-3 & relative to A-D-S\\
OSC type & sine,square,saw & tone type\\
IsNoise & boolean & whether to generate noise\\
Sample Length & 0-1 second & - \\
StartTime & 0-1 second & Length+Start<1\\
Amplitude & 0.1-1 & 1 = max amplitude\\
Pitches(notes) & list of pitches &  range of C0(16.35hz) to B9 \\
HP filter Cuttoff & 0-20000 & -\\
LP filter Cuttoff & 20000-HP & never lower than HP cutoff\\
Filter Order & 4,8,16 & butterworth filter order \\
\hline
\end{tabular}}
\caption{Synthesizer Sub-module Parameters. Despite the simplicity of the parameters and our efforts at constraining the ranges, the number of parameters that can be randomly chosen for each sub-module is in the order of $10^{15}$ }
\label{table:1}
\end{table}
\begin{figure*}[h]
\centering
\begin{subfigure}[b]{\linewidth}
\includegraphics[width=1\linewidth]{images/ff1.pdf}
\caption{Organic hat sample}\label{fig:64stack}
\setcounter{subfigure}{1}%
\includegraphics[width=1\linewidth]{images/ff2.pdf}
\caption{Randomly generated audio with percussive qualities, resembling a tight snare}\label{fig:64stack}
\setcounter{subfigure}{2}%
\includegraphics[width=1\linewidth]{images/ff3.pdf}
\caption{A randomly generated noise with a percussive envelop but non-percussive frequency features (modulated pitch)}\label{fig:64stack}
\setcounter{subfigure}{2}%
\end{subfigure}

@@

What we refer to as an "ear" is any method of evaluating a piece of audio, capable of "listening" to a piece of audio and giving it a score (or a list of scores) based on how well it satisfies certain criteria. Since in this work we are mainly focused on percussive generation, what we require from the ear is to give us probabilities of an audio sample belonging to various categories. As our synthesizer outputs are deterministic for all programs the ear's evaluations would allow us to associate the categorical probabilities of each sound with the program that generated it. In section \ref{gens} we discuss how these scores were used for navigation of our synthesis towards parameters which give us the desired sonic output.\\

@@

 \subsubsection{Survey of Randomly Generated Drums}
\label{survey}
\begin{figure}[H]
\centering
\includegraphics[width=1.1\linewidth]{images/cat.pdf}
\caption{Frequency of assigned labels for the human categorizers and the average categorizer}
\label{fig:freq-survey}
\end{figure}
To measure the quality of the samples produced by our pipeline and the power of our models, we randomly generated around 50-60 samples in the following categories: "snare", "kick", "hat", "clap" and "other" (combination of rims, 
shakers and unusual percussive sounds"). These samples were determined to be percussive and then categorized by 4 different models (FC, CNNLSTM, E+F and AVG). We ensured a balanced division between samples of stack size 1, 2 and 4 (each stack is responsible for a third of the samples under each category). Both reviewers then categorized these samples without knowledge of other categorizations (human or computation models). It's important to note that:
\begin{itemize}
    \item Humans had an additional category of "bad" for samples that they deemed not percussive. The "bad" category indicates that the sample should have been skipped in Phase 1. 
    \item With 6 categorization groups, humans had the same categorization in 47\% of cases.
    \item The agreement between the humans and AVG was 44\% and 47\%, not significantly lower than agreement with each other. 
    \item Of 257 samples humans agreed with FC, CNNLSTM, AVG and E+F respectively in 78, 76, 76 and 46 of cases.
\end{itemize}
We assess the reliability of agreement between humans and categorization models via the Fleiss' kappa coefficient \cite{fleiss1971measuring}. The value of 0 or less for this coefficient indicates no agreement beyond random chance, and the value of 1 indicating perfect agreement. Our kappa measurements \ref{kappa_table} lie within the 3.5-4.5 range, indicating mild to moderate agreement between humans and machines. We again measure this coefficient after dropping samples that were categorized as "bad" by the authors, as samples that humans deem to be "bad" indicate a failure in Phase 1 and arguably should not have been categorized by the models at all. Dropping of samples that both authors deemed "bad" causes an 8\% reduction of our data (21 samples) and a small increase in kappa score. Dropping samples either reviewer deemed bad resulted in a 30\% reduction of samples and relatively large increase in kappa scores. \\

@@

\begin{itemize}
    \item Our methodology was successful in generating novel examples of some percussion categories.
    \item Implementation of an ear that can distinguish sounds from closed sets (kick drums, snare drums etc)
    \item Our generation is not bullet-proof. Based on our human survay many of the samples generated in our pipeline are non-percussive. 
    \item There are potential benefits (e.g: pedagogical, software engineering, efficiency, data search, generality of solution, etc.) to separate some audio generation tasks into multiple parts, rather than throwing everything at some neural net. 
    \item With the right implementation of DSP methods to create a synth, Genetic search (viewed as too slow by previous work) is viable for finding desired programs and parameters.
    \item Neural nets are function approximators. You don't need them if the actual function is understandable and simple...(but is our task simple enough? how do I prove it?)
    \item When possible, using DSP methods instead of NNs is preferable in many ways (speed, computation, hardware requirements).\\
    \colorbox{green!=30}{Include some speed measurements in appendix?}
    \item ...
\end{itemize}
\section{Future Work}
\begin{itemize}
    \item Potential for other generative methods (e.g reinforcement learning)
    \item Making the synth's sound range more diverse while maintaining speed and efficiency
    \item Can our approach be used for VSTs?
\end{itemize}
